A Combinatorial Search of Parameterized Quantum Circuit Learning for Chemical Applications

### 2025-01-25 manuscript/main.tex
Sat Jan 25 12:16:35 PM EST 2025 manuscript BSE figure

Following model calibration using the function fitting dataset, we explore the appliciblity of PQCs for complex chemically relevant machine learning tasks.
The first chemically motivated dataset we explore is the BSE49 dataset, which contains the bond seperation energies (BSE)  for the homolytic bond cleavage of covalently bonded molecules, such as \ce{A-B -> A^{.} + B^{.}}.\cite{prasad_bse49_2021}
This dataset consists of 4394 datapoints, 1951 of which are existing and 2443 are hypthoetical structures, with 49 unique A-B single bond types.
In practice, we used 2436 of the hypothetical structures due to issues with valency exceptions when converting to RDKit mol objects which were later used for generating our features for the machine learning models.
An important aspect of machine learning in chemistry is the choice of molecular representation, or how the molecule is represented in the machine learning models.\cite{jones_molecular_2023}
Using RDKit\cite{noauthor_rdkit_nodate} we examined three commonly applied graph-based molecular preresentations, Molecular ACCess Systems (MACCS)\cite{durant_reoptimization_2002}, Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}, and RDKit fingerprints.
All three of these methods are use traversals of the molecular graphs to encode various structural details into bit vectors.
Lastly, we explore both topology- and physics-based molecular representations, both of which encode the three-dimensional structure of molecules in various, unique ways.
Persistent images (PIs) are a topology-based fingerprint that uses persistence homology to encode topological information of three-dimensional molecular structures into fixed dimension images.\cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022} 
We use the implementation from Townsend \textit{et al.}\cite{townsend_representation_2020}, which uses the Ripser Python package to generate PIs.\cite{tralie_ripserpy_2018}
Lastly, we explre two physics-based representations, Coulomb matrices (CMs) \cite{rupp_fast_2012} and smooth overlap of atomic positions (SOAPs), that were generated using DScribe.\cite{de_comparing_2016}
Due to the computational cost of computing the regularized entropy match (REMatch) kernel  with the SOAPs representation, we excluded this representation in the overall discussion.
We  also tested two different methods for representing the components of the bond separation chemical reaction, one where the feature vectors for the products are subtracted from the reactants, denoted by \textit{sub}, similar to the method used in Ref. \cite{garcia-andrade_barrier_2023}, and one that is composed of the reactant molecular only, denoted as \textit{AB}.
Since we are analyzing a diverse set of PQCs, we used a set of three different classic regression models, with varying capabilities, implemented in sckit-learn.\cite{pedregosa_scikit-learn_2011}
Based on our results shown in Fig. \ref{fig:classicalfuncfit} we found that Morgan fingerprints
We found that the best molecular representation across all models test, as shown in Fig. \ref{fig:classicalfuncfit}, was Morgan fingerprints using the \textit{sub} formulation.

\begin{figure}[H]
	\centering	
	\includegraphics[width=\linewidth]{../classical_testing/BSE/classical_funcfit}
	\caption{FIX:The performance of various molecular representations is shown using different classical machine learning models. \textit{AB} (top row) and \textit{sub} (bottom row) denote how the chemical reaction is represented. Each column denotes a different classical machine learning model, where the first column, denoted as knn, refers to \textit{k}-nearest neighbors, the second column, labeled rfr, denotes random forest regression, and krr, the last column, denotes kernel ridge regression. The x-axis of each sub-figure denotes the various molecular representations that are analyzed: Coulomb matrices (CM), Molecular ACCess Systems (MACCS), persistence images (PIs), RDKit fingerprints (RDKit), and Morgan fingerprints (Morgan).The y-axis of each sub-figure shows the performance of each model, evaluated using the R$^{2}$, where the training set is denoted using light blue and the test set is dark blue.}
	\label{fig:classicalfuncfit}
\end{figure}

>>>>

Following model calibration using the function fitting dataset, we explore the appliciblity of PQCs for complex chemically relevant machine learning tasks.
The first chemically motivated dataset we explore is the BSE49 dataset, which contains the bond seperation energies (BSE)  for the homolytic bond cleavage of covalently bonded molecules, such as \ce{A-B -> A^{.} + B^{.}}.\cite{prasad_bse49_2021}
This dataset consists of 4394 datapoints, 1951 of which are existing and 2443 are hypthoetical structures, with 49 unique A-B single bond types.
In practice, we used 2436 of the hypothetical structures due to issues with valency exceptions when converting to RDKit mol objects which were later used for generating our features for the machine learning models.
An important aspect of machine learning in chemistry is the choice of molecular representation, or how the molecule is represented in the machine learning models.\cite{jones_molecular_2023}
Using RDKit\cite{noauthor_rdkit_nodate} we examined three commonly applied graph-based molecular preresentations, Molecular ACCess Systems (MACCS)\cite{durant_reoptimization_2002}, Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}, and RDKit fingerprints.
All three of these methods are use traversals of the molecular graphs to encode various structural details into bit vectors.
Lastly, we explore both topology- and physics-based molecular representations, both of which encode the three-dimensional structure of molecules in various, unique ways.
Persistent images (PIs) are a topology-based fingerprint that uses persistence homology to encode topological information of three-dimensional molecular structures into fixed dimension images.\cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022} 
We use the implementation from Townsend \textit{et al.}\cite{townsend_representation_2020}, which uses the Ripser Python package to generate PIs.\cite{tralie_ripserpy_2018}
Lastly, we explre two physics-based representations, Coulomb matrices (CMs) \cite{rupp_fast_2012} and smooth overlap of atomic positions (SOAPs), that were generated using DScribe.\cite{de_comparing_2016}
Due to the computational cost of computing the regularized entropy match (REMatch) kernel  with the SOAPs representation, we excluded this representation in the overall discussion.
We  also tested two different methods for representing the components of the bond separation chemical reaction, one where the feature vectors for the products are subtracted from the reactants, denoted by \textit{sub}, similar to the method used in Ref. \cite{garcia-andrade_barrier_2023}, and one that is composed of the reactant molecular only, denoted as \textit{AB}.

[Ridge(), Lasso(), ElasticNet(), KNeighborsRegressor(), RandomForestRegressor(), GradientBoostingRegressor(), SVR(), KernelRidge(), GaussianProcessRegressor()]

Since we are analyzing a diverse set of PQCs, we used a set of three different classic regression models, with varying capabilities, implemented in sckit-learn.\cite{pedregosa_scikit-learn_2011}
Based on our results shown in Fig. \ref{fig:classical_molrepfig} we found that Morgan fingerprints
We found that the best molecular representation across all models test, as shown in Fig. \ref{fig:classical_molrepfig}, was Morgan fingerprints using the \textit{sub} formulation.

\begin{figure}[H]
	\centering	
	\includegraphics[width=\linewidth]{../images/BSE/classical/classical_molrepfig.png}
	\caption{}
	\label{fig:classical_molrepfig}
\end{figure}
